\large{Curtin University} \\ \large{Curtin Institute of Radio Astronomy} \\ \large{Physics Project 1} \\ \vspace*{20px} \textbf{\LARGE{Finding New Pulsars Using Machine Learning}}\vspace*{10px}

### 2020-06-03 Draft.tex
Abstract

>>>>

        As radio telescope technology improves, the number of candidate detections to be classified as pulsar or non-pulsar during pulsar surveys increases to the point that it becomes infeasible to classify them manually, a state of affairs known as the "pulsar candidate selection problem". To solve this problem, machine learning (ML) techniques are used to filter out the non-pulsar candidates in surveys, reducing the number of candidates requiring human evaluation to a manageable number. Here, we adapt an ML solution developed for the LOFAR Tied-Array All-Sky Survey (LOTAAS) for the Southern-sky Murchison Widefield Array Rapid Two-metre (SMART) survey. A group of 147 pulsar and 87 non-pulsar candidates generated by the SMART survey were used to test the LOTAAS ML classifiers trained on a set of 12 pulsar and 13 non-pulsar SMART survey candidates. The classifiers were found, on average, to make accurate pulsar and non-pulsar classifications on SMART survey candidates 87.3\% of the time, compared to the 97.1\% success rate obtained by \textcite{lyon}. Using the machine learning classifiers in ensemble with a selection criteria of three separate positive classifications increased the success rate of non-pulsar classification by 5.2\% (making 0 False Positive classifications), however it decreased the success rate of pulsar classification by 3.5\%. The ensemble machine learning classifier can be further improved by increasing the diversity of pulsars within the machine learning training dataset.
